Global code motion of congruent instructions

### 2017-09-11 paper.tex
typos

\citet{rosen1988global} also briefly discuss hoisting computations with
identical value numbers from immediate successors. Their algorithm iterates on
computations of same rank and move the code with identical computations from the
sibling branch to a common dominator if they are very busy
\cite{steven1997advanced}. The cost-model to mitigate register pressure is
missing, also there is no mention of sinking congruent instructions.

\GCC{} recently got code-hoisting \cite{GCCCodeHoisting} which is implemented as
part of \GVN{}-\PRE{}: it uses the set of ANTIC\_IN and AVAIL\_OUT value
expressions computed for \PRE{}. ANTIC\_IN[B] contains very busy expressions at
basic-block B, i.e., values computed on all paths from B to exit and
AVAIL\_OUT[B] contains values which are already available. The algorithm hoists
top down to a predecessor.  It uses ANTIC\_IN[B] to know what expressions will
be computed on every path from B to exit, and can be computed in B.  It uses
AVAIL\_OUT[B] to subtract out those values already being computed.  The cost
function is: for each hoist candidate, if all successors of B are dominated by
B, then we know insertion into B will eliminate all the remaining computations.
It then checks to see if at least one successor of B has the value available.
This avoids hoisting it way up the chain to ANTIC\_IN[B].  It also checks to
ensure that B has multiple successors, since hoisting in a straight line is
pointless.  The algorithm continues top down the dominator tree, working in
tandem with \PRE{} until no more hoisting is possible.  One advantage of \GCC{}
implementation is that it works in sync with the \GVN{}-\PRE{} such that when
new hoisting opportunities are created by \GVN{}-\PRE{}, code-hoisting will
hoist them.

@@

\citet{shobaki2013} also recently presented a combinatorial global
scheduler with reasonable performance improvements. It is possible that both
Barany and Shobaki's implementation will have similar results when compiled for
same target architecture. Also, both suffer from the same problem, although
Shobaki not so much, of large compile times which would not suit in industrial
compilers like \GCC{} and \LLVM{}. With the algorithm described in the next
section, we got a reduction in register spills, improved inliner heuristics,
improved compile time, and show performance improvements on \SPEC{} benchmarks.

@@

We compute merge sets, of the control flow graph, which remain same throughout
the \gcm{} transformation. For \gcm{} we only want to know if a use
operand is a kill (to compute changes in register pressure.) For that we only
need to know whether the use is also required later in the execution path. A
variable is live out of a basic-block B if it is used in the merge set of B. We
compute the live-out relation on-demand when profitability of hoist/sink
candidates is to be evaluated. A simplified version of
\emph{isLiveOutUsingMergeSet} is presented here:

@@

The \GCM{} pass can be broadly divided into the following steps that we will
describe in the rest of this section:
\begin{itemize}
\item find candidates (congruent instructions) suitable for \gcm{}
  (Section~\ref{subsec:finding-candidates},)
\item compute a point in the program where it is both legal
  (Section~\ref{subsec:legality}) and profitable
  (Section~\ref{subsec:cost-models}) to move the code,
\item move the code to hoist point or sink point
  (Section~\ref{subsec:optimistic},) and
\item update data structures to continue iterative \gcm{} (Section~\ref{subsec:optimistic}.)
\end{itemize}

>>>>

\citet{rosen1988global} also briefly discuss hoisting computations with
identical value numbers from immediate successors. Their algorithm iterates on
computations of same rank and move the code with identical computations from the
sibling branch to a common dominator if they are very busy
\cite{steven1997advanced}. However, the cost-model to mitigate register pressure
is missing, also, there is no mention of sinking congruent instructions.

\GCC{} recently got code-hoisting \cite{GCCCodeHoisting} which is implemented as
part of \GVN{}-\PRE{}: it uses the set of ANTIC\_IN and AVAIL\_OUT value
expressions computed for \PRE{}. ANTIC\_IN[B] contains very busy expressions
(VBEs) at basic-block B, i.e., values computed on all paths from B to exit and
AVAIL\_OUT[B] contains values which are already available. The algorithm hoists
VBEs to a predecessor.  It uses ANTIC\_IN[B] to know what expressions will be
computed on every path from the basic block B to exit, and can be computed in B.
It uses AVAIL\_OUT[B] to subtract out those values already being computed.  The
cost function is: for each hoist candidate, if all successors of B are dominated
by B, then we know insertion into B will eliminate all the remaining
computations.  It then checks to see if at least one successor of B has the
value available.  This avoids hoisting it way up the chain to ANTIC\_IN[B].  It
also checks to ensure that B has multiple successors, since hoisting in a
straight line is pointless.  The algorithm continues top down the dominator
tree, working in tandem with \PRE{} until no more hoisting is possible.  One
advantage of \GCC{} implementation is that it works in sync with the
\GVN{}-\PRE{} such that when new hoisting opportunities are created by
\GVN{}-\PRE{}, code-hoisting will hoist them.

@@

\citet{shobaki2013} also presented a combinatorial global scheduler with
reasonable performance improvements. It is possible that both Barany and
Shobaki's implementation will have similar results when compiled for same target
architecture. Also, both suffer from the same problem, although Shobaki not so
much, of large compile times which would not suit in industrial compilers like
\GCC{} and \LLVM{}. With the algorithm described in the next section, we got a
reduction in register spills, improved inliner heuristics, improved compile
time, and show performance improvements on \SPEC{} benchmarks.

@@

We compute merge sets, of the control flow graph, which remain same throughout
the \GCM{} transformation. For \GCM{} we only want to know if a use operand is a
kill (to compute changes in register pressure). For that we only need to know
whether the use is also required later in the execution path. A variable is live
out of a basic-block B if it is used in the merge set of B. We compute the
live-out relation on-demand when profitability of hoist/sink candidates is to be
evaluated. A simplified version of \emph{isLiveOutUsingMergeSet} is presented
here:

@@

The \GCM{} pass can be broadly divided into the following steps that we will
describe in the rest of this section:
\begin{itemize}
\item find candidates (congruent instructions) suitable for \gcm{}
  (Section~\ref{subsec:finding-candidates}),
\item compute a point in the program where it is both legal
  (Section~\ref{subsec:legality}) and profitable
  (Section~\ref{subsec:cost-models}) to move the code,
\item move the code to hoist point or sink point
  (Section~\ref{subsec:optimistic}) and
\item update data structures to continue iterative \gcm{} (Section~\ref{subsec:optimistic}).
\end{itemize}
